---
title: "Project1_README"
author: "Shashank P Shivdev"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Project 1

The data contains 332 comma-separated-value (CSV) files containing pollution monitoring data for fine particulate matter (PM) air pollution at 332 locations in the United States. Each file contains data from a single monitor and the ID number for each monitor is contained in the file name.

Each file contains three variables:
Date: the date of the observation in YYYY-MM-DD format (year-month-day)
sulfate: the level of sulfate PM in the air on that date (measured in micrograms per cubic meter)
nitrate: the level of nitrate PM in the air on that date (measured in micrograms per cubic meter)

### Part 1

To create a function that calculates the mean of a pollutant (sulfate or nitrate) across a specified list of monitors' ids, ignoring any missing values coded as NA.

```{r}

pollutantmean <- function(directory, pollutant, ids){
    
    setwd(directory)
    files = list.files(pattern = "*.csv") #list of csv files to be sourced
    
    masterlist <- list(NULL)
    index = 1
    
    for (file in files){
        df <- data.frame(read.csv(file = file))
        masterlist[[index]] <- df
        index = index + 1
    }
    
    rm(file, index, df) #remove intermediate variables
    
    meanlist = data.frame()
    
    for (i in ids){
        isna <- is.na(masterlist[[i]][ , pollutant])
        meanlist <- rbind(meanlist, masterlist[[i]][!isna, ])
    }
    
    mean(meanlist[ , pollutant])
    
}
```

### Part 2

To create a function that reads a directory full of csvs and reports the number of completely observed cases in each data file.

```{r}

complete <- function(directory, ids){
    
    setwd(directory)
    files = list.files(pattern = "*.csv") #list of csv files to be sourced
    
    masterlist <- list(NULL)
    index = 1
    
    for (file in files){
        df <- data.frame(read.csv(file = file))
        masterlist[[index]] <- df
        index = index + 1
    }
    
    rm(file, index, df) #remove intermediate variables
    
    complete_list <- data.frame()
    
    for (i in ids){
        nobs <- nrow(masterlist[[i]][complete.cases(masterlist[[i]]), ])
        complete_list <- rbind(complete_list, c(i, nobs))
    }
    
    colnames(complete_list) <- c("id", "nobs")
    complete_list
}
```

### Part 3

To create a function that takes a directory of data files and a threshold for complete cases and calculates the correlation between sulfate and nitrate for monitor locations where the number of completely observed cases (on all variables) is greater than the threshold.

```{r}
corr <- function(directory, threshold = 0){
    
    setwd(directory)
    files = list.files(pattern = "*.csv") #list of csv files to be sourced
    
    masterlist <- list(NULL)
    index = 1
    
    for (file in files){
        df <- data.frame(read.csv(file = file))
        masterlist[[index]] <- df
        index = index + 1
    }

    
    rm(file, index, df) #remove intermediate variables
    
    corr_vect <- NULL
    
    for (i in 1:length(masterlist)){
        nobs <- nrow(masterlist[[i]][complete.cases(masterlist[[i]]), ])
        if(nobs > threshold){
            corr_vect <- c(corr_vect, cor(masterlist[[i]][complete.cases(masterlist[[i]]), "sulfate"], masterlist[[i]][complete.cases(masterlist[[i]]), "nitrate"]))
        }
    }
    
    corr_vect
}

```

