---
title: "Project3_README"
author: "Shashank P Shivdev"
date: "November 3, 2017"

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Project 3

The goal is to prepare tidy data that can be used for later analysis.

The data linked represents data collected from the accelerometers from the Samsung Galaxy S smartphone. A full description is available at the site where the data was obtained:

http://archive.ics.uci.edu/ml/datasets/Human+Activity+Recognition+Using+Smartphones

Here are the data for the project:

https://d396qusza40orc.cloudfront.net/getdata%2Fprojectfiles%2FUCI%20HAR%20Dataset.zip

### Data Processing - Steps

#### Step 1

Downloading and unzipping dataset

```{r}

if(!file.exists("./data")){dir.create("./data")}
fileUrl <- "https://d396qusza40orc.cloudfront.net/getdata%2Fprojectfiles%2FUCI%20HAR%20Dataset.zip"
download.file(fileUrl,destfile="./data/Dataset.zip")

unzip(zipfile="./data/Dataset.zip",exdir="./data")

```

#### Step 2

Merging the training and the test sets to create one data set:
Reading files

```{r}

features <- read.table("E:/DS_C/R_projects/3/UCI_data/features.txt")

activityLabels <- read.table("E:/DS_C/R_projects/3/UCI_data/activity_labels.txt", col.names = c('activityId','activityType'))

x_train <- read.table("E:/DS_C/R_projects/3/UCI_data/train/X_train.txt", col.names = features[ , 2])
y_train <- read.table("E:/DS_C/R_projects/3/UCI_data/train/y_train.txt", col.names = "activityId")
subject_train <- read.table("E:/DS_C/R_projects/3/UCI_data/train/subject_train.txt", col.names = "subjectId")

x_test <- read.table("E:/DS_C/R_projects/3/UCI_data/test/X_test.txt", col.names = features[ , 2])
y_test <- read.table("E:/DS_C/R_projects/3/UCI_data/test/y_test.txt", col.names = "activityId")
subject_test <- read.table("E:/DS_C/R_projects/3/UCI_data/test/subject_test.txt", col.names = "subjectId")

```

Merging datasets

```{r}

total_train <- cbind(y_train, subject_train, x_train)
total_test <- cbind(y_test, subject_test, x_test)
total_data <- rbind(total_train, total_test)

```

#### Step 3

Extracting only the measurements on the mean and standard deviation for each measurement

```{r}

colNames <- colnames(total_data)

sdmu_colNames <- (grepl("activityId" , colNames) | grepl("subjectId" , colNames) | grepl("mean.." , colNames) | grepl("std.." , colNames))

sdmu_data <- total_train[ , sdmu_colNames == TRUE]

```

#### Step 4

Using descriptive activity names to name the activities in the data set

```{r}

activitynames_data <- merge(sdmu_data, activityLabels, by = 'activityId', all = TRUE)

```

#### Step 5

Creating a second, independent tidy data set with the average of each variable for each activity and each subject

```{r}

tidyset <- aggregate(. ~subjectId + activityId, activitynames_data, mean)
tidyset <- tidyset[order(tidyset$subjectId, tidyset$activityId), ]
```

Writing second tidy data set in txt file

```{r}

write.table(tidyset, "tidyset.txt", row.name=FALSE)