---
title: "Project2_README"
author: "Shashank P Shivdev"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Project 2

The data contains two files - 
1. Outcome-of-care-measures.csv: Contains information about 30-day mortality and readmission rates for heart attacks, heart failure, and pneumonia for over 4,000 hospitals. 
2. Hospital-data.csv: Contains information about each hospital. 

### Part 1

To create a function that returns the name of the hospital that has the best (i.e. lowest) 30-day mortality for the specified outcome in that state.

```{r}

best<- function(state, outcome){
    
  data <- read.csv("outcome-of-care-measures.csv", colClasses = "character")
  
  if(!any(state == data$State)){
    stop(print("Invalid state"))
    }
  else if((outcome %in% c("heart attack", "heart failure", "pneumonia")) == FALSE){
    stop(print("Invalid outcome"))
    }
  
  sub_data <- subset(data, State == state)
  if (outcome == "heart attack"){
    colnum <- 11
    }
  else if (outcome == "heart failure"){
    colnum <- 17
    }
  else {
    colnum <- 23
    }
  
  min <- which(as.numeric(sub_data[ ,colnum]) == min(as.numeric(sub_data[ ,colnum]), na.rm = TRUE))
  hospitals <- sub_data[min, 2]
  hospitals <- sort(hospitals)
  
  return(hospitals[1])
  
}
```

### Part 2

To create a function that reads the outcome-of-care-measures.csv file and returns the name of the hospital that has the ranking specified by an argument for a particular specified outcome

```{r}

best<- function(state, outcome, num){
    
  data <- read.csv("outcome-of-care-measures.csv", colClasses = "character")
  
  if(!any(state == data$State)){
    stop(print("Invalid state"))
    }
  else if((outcome %in% c("heart attack", "heart failure", "pneumonia")) == FALSE){
    stop(print("Invalid outcome"))
    }
  
  sub_data <- subset(data, State == state)
  if (outcome == "heart attack"){
    colnum <- 11
    }
  else if (outcome == "heart failure"){
    colnum <- 17
    }
  else{
    colnum <- 23
    }
  
  sub_data[, colnum] <- as.numeric(sub_data[, colnum])
  sub_data <- sub_data[order(sub_data[, colnum], sub_data[, 2]), ]
  sub_data <- sub_data[(!is.na(sub_data[, colnum])), ]
  
  if(num == "best"){
    num <- 1
    }            
  else if (num == "worst"){
    num <- nrow(sub_data)
    }      
  
  return(sub_data[num, 2])

}
```

### Part 3

To create a function that reads returns a 2-column data frame containing the hospital in each state that has the ranking specified for a specified outcome

```{r}

rankall<- function(outcome, num = "best"){
    
     data <- read.csv("outcome-of-care-measures.csv", colClasses = "character")
    
     if((outcome %in% c("heart attack", "heart failure", "pneumonia")) == FALSE){
         stop(print("invalid outcome"))
    }
    
     if (outcome == "heart attack"){
        colnum <- 11
    }
    else if (outcome == "heart failure"){
        colnum <- 17
    }
    else{
        colnum <- 23
    }
    
    data[, colnum] <- as.numeric(data[, colnum])
    
    data = data[!is.na(data[, colnum]), ]
    
    split_data = split(data, data$State)
    
    result = lapply(split_data, function(x, num){ 
        
        x = x[order(x[, colnum], x$Hospital.Name), ]
      
        if(num == "best"){
            return (x$Hospital.Name[1])
        }
        else if(num == "worst"){
            return (x$Hospital.Name[nrow(x)])
        }
        else{
            return (x$Hospital.Name[num])
        }
    }, num)
    
    return(data.frame(Hospital=unlist(result), State=names(result)))
}
```
